---
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
dir.create("images")
```



```{r}
#install.packages
library(caret)
library(ISLR)
library(gmodels)
library(ggplot2)
library(FNN)
library(dummies)
library(fastDummies)
library(e1071)
library(dplyr)

#install.packages("caret")
library(caret)
# install.packages("ISLR") # only install if needed
library(ISLR)
```


```{r}
library(readxl)
uk <- read.csv("C:/Users/mnooo/Desktop/Datasets/UniversalBank.csv")
#View(uk)

```



```{r}

# Clean up data and get it ready for KNN
library(dplyr)


uk$Education_1 = ifelse(uk$Education ==1, 1, 0)
uk$Education_2 = ifelse(uk$Education ==2, 1, 0)
uk$Education_3 = ifelse(uk$Education ==3, 1, 0)
uk$Personal_Loan = as.factor(uk$Personal.Loan)

View(uk)

uk1 <- uk[, c(2:4, 6:7, 9, 11:18)]   # Select a subset of variables, without ID and zipcode, original variable Personal.Loan
head(uk1)
str(uk1)
```


```{r}
set.seed(15)
Test_Index = createDataPartition(uk1$Age,p=0.2, list=FALSE) # 20% reserved for Test
Test_Data = uk1[Test_Index,]
TraVal_Data = uk1[-Test_Index,] # Validation and Training data is rest
Train_Index = createDataPartition(TraVal_Data$Age,p=0.75, list=FALSE) # 75% of remaining data as training

# 75% of the 80% is 60% which is training

Train_Data = TraVal_Data[Train_Index,]
Validation_Data = TraVal_Data[-Train_Index,] # rest as validation
View(Train_Data)
summary(Train_Data)
summary(Validation_Data)
summary(Test_Data)
str(Train_Data)
str(Validation_Data)
str(Test_Data)

```
```{r}
#Normalization

# Copy the original data
train.norm.df <- Train_Data
valid.norm.df <- Validation_Data
traval.norm.df <- TraVal_Data
# use preProcess() from the caret package
norm.values <- preProcess(Train_Data[, 1:13], method=c("center", "scale"))
train.norm.df[, 1:13] <- predict(norm.values, Train_Data[, 1:13]) # Replace first two columns with normalized values
valid.norm.df[, 1:13] <- predict(norm.values, Validation_Data[, 1:13])
traval.norm.df[, 1:13] <- predict(norm.values, traval.norm.df[, 1:13])
test.norm.df <- predict(norm.values, Test_Data[, 1:13])
summary(train.norm.df)
var(train.norm.df[, 1:13])
summary(valid.norm.df)
var(valid.norm.df[, 1:13])

```


```{r}

dim(train.norm.df[, 1:13]) # Train is 13 numeric variable 
dim(test.norm.df)         # Test is 13 numeric variable 
table(train.norm.df[, 14])  # DV is logic (Y/N) variable
library(FNN)
nn <- knn(train = train.norm.df[, 1:13], test = test.norm.df, 
          cl = train.norm.df[, 14], k = 1, prob=TRUE) # We use k = #Q1

customer_df<- data.frame("Age" =40, "Experience"=10, "Income"=84, "Family"=2, "CCAvg"=2, "Mortgage"=0,  "Securities Account"=0, "CD Account"=0,  "Online" =1, "Credit Card"=1, "Education_1" = 0, "Education_2" = 1, "Education_3" = 0)

customerClass <- knn(train.norm.df[, 1:13], customer_df,  cl = train.norm.df[, 14], k = 1, prob = 0.5)

paste0("The customer is classified as: ", summary(customerClass))

# print(nn) # uncomment for more output
# row.names(Train_Data)[attr(nn, "nn.index")]

#table(nn)
#table(train.norm.df[, 14])
#table(valid.norm.df[, 14])

#nn <- knn(train.norm.df[, 1:13], valid.norm.df[, 1:13], cl = train.norm.df[, 14], k = 1)

#confusionMatrix(nn, valid.norm.df[, 14])$overall[1] 

```
Q1 Answer: The customer is classified as: 1 (loan acceptance)
```{r}
#Q2
# initialize a data frame with two columns: k, and accuracy.
library(caret)
accuracy.df <- data.frame(k = seq(1, 14, 1), accuracy = rep(0, 14))
# compute knn for different k on validation.
for(i in 1:14) {
  knn.pred <- knn(train.norm.df[, 1:13], valid.norm.df[, 1:13], 
                  cl = train.norm.df[, 14], k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.pred, valid.norm.df[, 14])$overall[1] 
}
accuracy.df
```
The value of k that provides the best performance is k = 3. 
k=3 gives the highest accuracy percentage of 95.5% however it should be noted that this classification model is built based on actual converted customers percentage of only 9.8% of the whole dataset which might be causing overfitting in the model

```{r}
#Q3 - Confusion Matrix 

knn.pred <- knn(train.norm.df[, 1:13], valid.norm.df[, 1:13], 
                  cl = train.norm.df[, 14], k = 3)
confusionMatrix(knn.pred, valid.norm.df[, 14])$overall[1]
confusionMatrix(knn.pred, valid.norm.df[, 14])

```
```{r}
#Q4
customerClass <- knn(train.norm.df[, 1:13], customer_df,  cl = train.norm.df[, 14], k = 3, prob = 0.5)

paste0("The customer is classified as: ", summary(customerClass))

```
 The customer is classified as: 1 (loan acceptance)

```{r}
#Q5
set.seed(15)
Test_Index = createDataPartition(uk1$Age,p=0.2, list=FALSE) # 20% reserved for Test
Test_Data = uk1[Test_Index,]
TraVal_Data = uk1[-Test_Index,] # Validation and Training data is rest
Train_Index = createDataPartition(TraVal_Data$Age, p=0.625, list=FALSE) # 62.5% of remaining data as training

# 62.5% of the 80% is 50% which is training

Train_Data = TraVal_Data[Train_Index,]
Validation_Data = TraVal_Data[-Train_Index,] # rest as validation
```

 ## Normalization
 
```{r}


# Copy the original data
train.norm.df <- Train_Data
valid.norm.df <- Validation_Data
traval.norm.df <- TraVal_Data
# use preProcess() from the caret package to normalize Sales and Age.
norm.values <- preProcess(Train_Data[, 1:13], method=c("center", "scale"))
train.norm.df[, 1:13] <- predict(norm.values, Train_Data[, 1:13]) # Replace first two columns with normalized values
valid.norm.df[, 1:13] <- predict(norm.values, Validation_Data[, 1:13])
traval.norm.df[, 1:13] <- predict(norm.values, traval.norm.df[, 1:13])
test.norm.df <- predict(norm.values, Test_Data[, 1:13])
summary(train.norm.df)
var(train.norm.df[, 1:13])
summary(valid.norm.df)
var(valid.norm.df[, 1:13])

```

  ## KNN
```{r}

library(FNN)
nn <- knn(train = train.norm.df[, 1:13], test = test.norm.df, 
          cl = train.norm.df[, 14], k = 3, prob=TRUE) 
```

  ## Confusion matrix

```{r}

knn.pred.valid <- knn(train.norm.df[, 1:13], valid.norm.df[, 1:13], 
                  cl = train.norm.df[, 14], k = 3)
confusionMatrix(knn.pred.valid, valid.norm.df[, 14])
```

```{r}

knn.pred.train <- knn(train.norm.df[, 1:13], train.norm.df[, 1:13], 
                  cl = train.norm.df[, 14], k = 3)
confusionMatrix(knn.pred.train, train.norm.df[, 14])
```



