---
title: "Assignment 3"
author: "MUNERAH"
date: "10/11/2021"
output: html_document
---

```{r}
library(readxl)
df<- read.csv("C:/Users/mnooo/Desktop/Datasets/UniversalBank.csv")
str(df)
```


```{r }
#install.packages
library(dplyr)
library(class)
library(caret)
library(FNN)
library(e1071)
library(reshape2)
```



```{r}
### Change Numerical data to Catogerical 
df$Personal.Loan<-factor(df$Personal.Loan)
df$Online<-factor(df$Online)
df$CreditCard <-factor(df$CreditCard)
str(df)

```

```{r}
### Divide the data into 60% training and 40% validation
# First select the required variables
selected.var <- c(10,13,14)
set.seed(15)
bank.tr.in = createDataPartition(df$Personal.Loan,p=0.6, list=FALSE) # 60% reserved for Training
bank.tr = df[bank.tr.in,selected.var]
bank.va <- df[-bank.tr.in,selected.var] # Validation  data is rest
summary(bank.tr)
summary (bank.va)

```

```{r}
###Create a pivot table for the training data with Online as a column variable, CC as a row variable, and Loan as a secondary row variable.

### using table function
table(bank.tr)
```



```{r}
##B. Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services. Looking at the pivot table, what is the probability that this customer will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1)].


P1<-prop.table(table(bank.tr),margin = 2)
P1
```
Looking to the pivot table , There are 507 customers who owns a bank credit card and actively using online service , 46 of them will accept the loan 
The probability of a customer to accept the loan conditional on having a credit card and being an active of online services  is (46/507)*100 = 9.07%
In another way : 0.0261/(0.0261+0.261) = 0.09


```{r}
##C. Create two separate pivot tables for the training data. One will have Loan (rows) as a function of Online (columns) and the other will have Loan (rows) as a function of CC.

Pivot1 <- table(bank.tr$Personal.Loan,bank.tr$Online)
pivot_df1 <- as.data.frame(Pivot1)
colnames(pivot_df1) <- c("PersonalLoan", "Online")
pivot_df1

Pivot2 <- table(bank.tr$Personal.Loan,bank.tr$CreditCard)
pivot_df2 <- as.data.frame(Pivot2)
colnames(pivot_df2) <- c("PersonalLoan", "CreditCard")
pivot_df2






```
```{r}
##D. Compute the following quantities [P(A | B) means “the probability of A given B

##i. P(CC = 1 | Loan = 1) 

Pr1<- (table(bank.tr$CreditCard,bank.tr$Personal.Loan))
Pr1[2,2]/(Pr1[2,2]+Pr1[1,2])
Pr1



```

The result as shown is 28%

```{r}

##ii. P(Online = 1 | Loan = 1)

Pr2<-table(bank.tr$Online , bank.tr$Personal.Loan)
Pr2[2,2]/(Pr2[2,2]+Pr2[1,2])
Pr2

```

The result as shown 59%
```{r}
##iii.P(Loan = 1) (the proportion of loan acceptors)
Pr3<-table(bank.tr$Personal.Loan)
Pr3[2]/(Pr3[2]+Pr3[1])
Pr3



```
The result as shown approx 10%
```{r}
##iv. P(CC = 1 | Loan = 0)
Pr4<-table(bank.tr$CreditCard,bank.tr$Personal.Loan)
Pr4[2,1]/(Pr4[2,1]+Pr4[1,1])
Pr4



```
The result as shown is 29%


```{r}

##v.P(Online = 1 | Loan = 0)
Pr5<-table(bank.tr$Online , bank.tr$Personal.Loan)
Pr5[2,1]/(Pr4[2,1]+Pr4[1,1])
Pr5


```
The result as shown is 59%
```{r}
## vi. P(Loan = 0)
Pr6<-table(bank.tr$Personal.Loan)
Pr6[1]/(Pr5[1]+Pr5[2])
Pr6



```
The result as shown approx 90%
```{r}
##E. Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC = 1, Online = 1)
### C1 loan , X1 CC ,X2 Online 
###P(C1|X1,X2) = 
##(P(X1|C1)P(X2|C1)P(C1) ) / P(X1|C1)P(X2|C1)P(C1)+P(X1|C2)*P(X2|C2)*P(C2)
x1<-((0.28125)*(0.593)*(0.096)) 
x2<-((0.28125)*(0.593)*(0.096))+((0.2916)*(0.5866)*(0.904))
nbresult<- (x1/x2)
nbresult

```
```{r}
##F. Compare this value with the one obtained from the pivot table in (B). Which is a more accurate estimate?

# The result in B = 0.0907 ,in E =0.938,, there is no big different between them , and the value which is calculated by pivot table is more accurate because the Naive Base assume the probabilities being independent 



```
```{r}
##G. Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)? Run
##naive Bayes on the data. Examine the model output on training data, and find the entry that
##corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E).

bank.nb <- naiveBayes(Personal.Loan ~ ., data = bank.tr)
bank.nb


```
```{r}((0.28125)*(0.593)*(0.096)) /
((0.28125)*(0.593)*(0.096))+((0.2916)*(0.5866)*(0.904))
= 0.0938
##it gives identical result as it is in the question E . 

```